Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae

### 2021-05-10 thesis_chapter_3.Rmd
new corrections and ref error sorted

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, 41 Market Place Viaduct Harbour, 1010 Auckland, Private Bag 99940, Newmarket, Auckland 1149, New Zealand], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
output:
  word_document:
    reference_docx: markdown/wiley_template.docx
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document:
bibliography: references.bib
csl: markdown/limnologyandoceanography.csl
abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalgae are one of the most important natural dispersal vectors  of marine organisms and are regarded as the 'tumble-weed of the ocean'. Despite many studies that exist on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still unknown. Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a consenses as to the relative role of wind versus surface currents. In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam. Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation. We focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddys within the study domain. The results also point to the importance of sub-mesoscale features, such as eddys, as important surface accumulation zones for solitary floating macroalage."
---

@@

The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996]. The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324 km [@lutjeharms1988;@lutjeharms2006; @lutjeharms2007]. Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015]. The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009]. The Benguela region ---known as the Benguela Upwelling System (BUS)--- is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017]. The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996]. The region is also characterized by large amounts of mesoscale variability in the form of eddys and filaments [@blanke2002; @blanke2005; @rubio2009]. Inshore, the BUS is dominated by a northwestward flow, known as the Benguela Jet, because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddys [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009]. The stability of the jet varies seasonally with more stable flows in winter; in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch. The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009]. These winds are most intense during Austral summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019]. In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].

@@

Virtual particles were released with no drag or wind effects from the study site to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag and are referred to hereafter as 'passive particles'. All particles that have any form of drag exposure are hereafter referred to as 'kelp particles'. All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018]. Brownian motion was also incorporated to account for any stochasticity not resolved by the model. These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018]. The time period used in all simulations were January (Austral summer), April (Austral autumn), July (Austral winter), and October (Austral spring) of 2018. These time periods were chosen based on variable mesoscale activities and stability of the Bengulea Jet that occurs through the year. For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity while in winter the Benguela Jet begins to stabilize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].

@@

## Model inputs
### Ocean and wind model

@@

### Hydrodynamic and wind drag

@@

waterdrag_plot <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
  geom_path(data = total_mean_water, aes(x = lon, y = lat, linetype = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

wwdrag_plot <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
  geom_path(data = total_mean_ww, aes(x = lon, y = lat, linetype = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

@@

waterdrag_plot_April <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
  geom_path(data = total_mean_water_April, aes(x = lon, y = lat, linetype = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

wwdrag_plot_April <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill ="grey") +
  geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, linetype = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

@@

waterdrag_plot_July <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
  geom_path(data = total_mean_water_July, aes(x = lon, y = lat, linetype = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

wwdrag_plot_July <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
  geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, linetype = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

@@

waterdrag_plot_Oct <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
  geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, linetype = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(linetype=guide_legend(title="Simulation")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme_bw()

wwdrag_plot_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
  geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, linetype = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(linetype=guide_legend(title="Simulation")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme_bw()

@@

Across all simulations, particles flowed in a north-westward direction. The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids. In addition, the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 4). Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.

In April and across all simulations, particles flowed in a north-westward direction (Figure 4). The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation. In addition, simulations that considered any form of drag got entrained within an eddy further north-westward from the end location of passive particles. Only slight variations across cross-sectional areas and drag exposure scenarios are evident. Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (Figure 5). Variation in density along the mean trajectory and eddy for all kelp particles is evident. Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag. Simulations run for October show no difference in density of particles along the mean trajectory when comparing passive and kelp particles. Only slight variations in density and trajectory can be seen with no entrainment (Figure 6). Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.

@@

heatmap_RK4 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
  geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H100_W00 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H100W100") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H095_W05 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  #ggtitle("H095W05") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H090_W10 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H090W10") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H085_W15 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
  geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H085W15") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

>>>>

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
output:
  word_document:
    reference_docx: markdown/wiley_template.docx
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document:
bibliography: references.bib
csl: markdown/limnologyandoceanography.csl
abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalgae are one of the most important natural dispersal vectors  of marine organisms and are regarded as the 'tumble-weed of the ocean'. Despite many studies that exist on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still unknown. Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a consenses as to the relative role of wind versus surface currents. In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam. Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation. We focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddys within the study domain. The results also point to the importance of sub-mesoscale features, such as eddys, as important surface accumulation zones for solitary floating macroalage."
---

@@

The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996]. The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324 km [@lutjeharms1988;@lutjeharms2006; @lutjeharms2007]. Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015]. The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009]. The Benguela region ---known as the Benguela Upwelling System (BUS)--- is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017]. The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996]. The region is also characterized by large amounts of mesoscale variability in the form of eddys and filaments [@blanke2002; @blanke2005; @rubio2009]. Inshore, the BUS is dominated by a northwestward flow, known as the Benguela Jet, because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddys [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009]. The stability of the jet varies seasonally with more stable flows in winter; in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch. The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009]. These winds are most intense during Austral summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019]. In Austral winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].

@@

Virtual particles were released with no drag or wind effects from the study site to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag and are referred to hereafter as 'passive particles'. All particles that have any form of drag exposure are hereafter referred to as 'kelp particles'. All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @vanSebille2018]. Brownian motion was also incorporated to account for any stochasticity not resolved by the model. These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019;@vanSebille2018]. The time period used in all simulations were January (Austral summer), April (Austral autumn), July (Austral winter), and October (Austral spring) of 2018. These time periods were chosen based on variable mesoscale activities and stability of the Bengulea Jet that occurs through the year. For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity while in winter the Benguela Jet begins to stabilize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].

@@

## Model inputs
## Ocean and wind model

@@

## Hydrodynamic and wind drag

@@

waterdrag_plot <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_path(data = total_mean_water, aes(x = lon, y = lat, linetype = type))+
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

wwdrag_plot <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_path(data = total_mean_ww, aes(x = lon, y = lat, linetype = type))+
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

@@

waterdrag_plot_April <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_path(data = total_mean_water_April, aes(x = lon, y = lat, linetype = type))+
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

wwdrag_plot_April <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill ="grey") +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, linetype = type))+
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

@@

waterdrag_plot_July <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_path(data = total_mean_water_July, aes(x = lon, y = lat, linetype = type))+
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

wwdrag_plot_July <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, linetype = type))+
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

@@

waterdrag_plot_Oct <- ggplot()+
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, linetype = type))+
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(linetype=guide_legend(title="Simulation")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

wwdrag_plot_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, linetype = type))+
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(linetype=guide_legend(title="Simulation")) +
  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
  theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12)) 
```

@@

Across all simulations for January (Austral summer), particles flowed in a north-westward direction. The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 4). In addition, the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 4). Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.

In April and across all simulations, particles flowed in a north-westward direction (Figure 5). The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation. In addition, simulations that considered any form of drag got entrained within an eddy further north-westward from the end location of passive particles. Only slight variations across cross-sectional areas and drag exposure scenarios are evident. Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (Figure 6). Variation in density along the mean trajectory and eddy for all kelp particles is evident. Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag. Simulations run for October show no difference in density of particles along the mean trajectory when comparing passive and kelp particles. Only slight variations in density and trajectory can be seen with no entrainment (Figure 7). Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.

@@

load("map/south_africa_coast_hi_res.Rdata")
load("map/sa_bathy.RData")

heatmap_RK4 <- ggplot() +
  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 0.4, colour = "black", fill = "gray") +
  geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  ggtitle("1") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

heatmap_Min_Kfloat_H100_W00 <- ggplot() +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis",limits = c(1, 1100))  +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 0.4, colour = "black", fill = "gray") +
  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  ggtitle("2") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

heatmap_Min_Kfloat_H095_W05 <- ggplot() +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 0.4, colour = "black", fill = "gray") +
  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  guides(fill=guide_legend(title="Number of particles")) +
  ggtitle("3") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))

heatmap_Min_Kfloat_H090_W10 <- ggplot() +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 0.4, colour = "black", fill = "gray") +
  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  labs(x = expression(Longitude*degree*S)) +
  labs(y = expression(Latitude*degree*E)) +
  ggtitle("4") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12))
